---
title: "DiD Course Problem Set 2"
author: "Pedro Scatimburgo"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: readable
    toc: true
    number_sections: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
```

<style type="text/css">

h1.title {
  font-size: 24px;
  color: black;
  text-align: center;
}
subtitle {
  font-size: 24px;
  color: black;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 16px;
  font-family: "Kaleko205Text-Bold", Kaleko 205, Bold;
  color: black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 16px;
  font-family: "Kaleko205Text-Bold", Kaleko 205, Bold;
  color: black;
  text-align: center;
}
</style>

# Problem Set 1

## Define potential outcomes depending on potential treatment sequences.

$Y_{i,1}(d_1)$, $Y_{i,2}(d_1,d_2)$, $Y_{i,3}(d_1,d_2,d_3)$.

## Define the average treatment effect at time $t$ of taking a specific treatment sequence compared to never being treated, among units that take that given specified treatment sequence.

Consider a specific treatment $d^\prime$. Let $\vec{0} = (0,0,0)$, $d^\prime = (d_1',d_2',d_3')$. Then:

\begin{equation}
ATE(d',\vec{0},t) = \mathbf{E}\left[Y_{i,t}(d')-Y_{i,t}(\vec{0})|d=d'\right]
\end{equation}

## Define the average treatment effect at time $t$ of taking a specific treatment sequence compared to never being treated, among units that remained untreated until $t=3$.

\begin{equation}
ATE(d',\vec{0},t) = \mathbf{E}\left[Y_{i,t}(d')-Y_{i,t}(\vec{0})|d=(0,0,1)\right]
\end{equation}

## Define the overall average treatment effect at time $t$ of taking a specific treatment sequence compared to never being treated.

\begin{equation}
ATE(d',\vec{0},t) = \mathbf{E}\left[Y_{i,t}(d')-Y_{i,t}(\vec{0})\right]
\end{equation}

## Define potential outcomes depending on treatment dosages.

$Y_{i,t}(d_1)$, $Y_{i,2}(d_1,d_2)$, where $d = (d_1,d_2) >> 0$.

## Define the overall average treatment effect at time $t=2$ of receiving dosage $d$ versus not receiving any treatment.

\begin{equation}
ATE(d,0,2) = \mathbf{E}\left[Y_{i,2}(0,d)-Y_{i,2}(0,0)\right]
\end{equation}

## Define the overall average treatment effect at time $t=2$ of receiving dosage $d$ versus receiving dosage $d^\prime$.

\begin{equation}
ATE(d,d',2) = \mathbf{E}\left[Y_{i,2}(0,d)-Y_{i,2}(0,d')\right]
\end{equation}

## Define the average treatment effect at time $t=2$ of receiving dosage $d$ versus not receiving any treatment, among units who received dosage $d$.

\begin{equation}
ATE(d,0,2) = \mathbf{E}\left[Y_{i,2}(0,d)-Y_{i,2}(0,0)|d^*=d>0\right]
\end{equation}

## Define the average treatment effect at time $t=2$ of marginally increasing treatment dosage $d$, among units who received dosage $d$. Discuss the discrete and continuous cases separately.

In the discrete case:

\begin{equation}

ATE(d+\varepsilon,d,2) = \mathbb{E}\left[Y_{i,2}(0,d+\varepsilon)-Y_{i,2}(0,d)|d^*=d>0\right]

\end{equation}

In the continuous case:

\begin{equation}

ATE = \mathbb{E}\left[\left.\frac{\partial Y_{i,t}(0,d)}{\partial d} \right\vert d^*=d>0 \right]

\end{equation}

## Define the overall average treatment effect at time $t=2$ of marginally increasing treatment dosage $d$. Discuss the discrete and continuous case separately.

In the discrete case:

\begin{equation}

ATE(d+\varepsilon,d,2) = \mathbb{E}\left[Y_{i,2}(0,d+\varepsilon)-Y_{i,2}(0,d)\right]

\end{equation}

In the continuous case:

\begin{equation}

ATE = \mathbb{E}\left[\frac{\partial Y_{i,t}(0,d)}{\partial d}\right]

\end{equation}


## The above marginal average treatment effects are “local“ to a dosage $d$. Can you think of a more aggregate treatment effect measure that may summarize the above marginal average treatment effects across different dosages $d$?

# Problem Set 2

## Load the dataset in your statistical software of choice.

```{r}
data <- haven::read_dta("data/ehec_data.dta")
```


## For simplicity, let’s subset the data such that we have only 2 time periods, ``year = 2013`` and ``year = 2014``.

```{r}
data <- data %>%
  filter(year %in% c("2013","2014"))
```


## Ensure that ``stfips`` and ``year`` are numerical variables (this may matter in some software like R).

```{r}
data <- data %>%
  mutate(
    year = as.numeric(as.character(year)),
    stfips = as.numeric(as.character(stfips))
  )
```


## Create a treatment group dummy that takes value 1 if a state expanded Medicaid by 2014, and 0 otherwise. Also create a post-treatment period dummy that takes value 1 if ``year = 2014`` and 0 otherwise.

```{r}
data <- data %>%
  mutate(
    treatment = replace_na(ifelse(yexp2 == 2014, 1, 0), 0),
    post_treatment = ifelse(year == 2014, 1, 0),
    product = treatment*post_treatment
  )
```


## Compute the DiD estimator for the ATT “by-hand”, i.e., using simple comparisons of means. Repeat the same exercise by using W as weights.

```{r}
# Without using W as weights:

first_diff <- mean(data %>% filter(treatment == 1 & post_treatment == 1) %>% pull(dins)) - mean(data %>% filter(treatment == 1 & post_treatment == 0) %>% pull(dins))

second_diff <- mean(data %>% filter(treatment == 0 & post_treatment == 1) %>% pull(dins)) - mean(data %>% filter(treatment == 0 & post_treatment == 0) %>% pull(dins))

(did_estimator <- first_diff - second_diff)

# Using W as weights:

first_diff_W <- weighted.mean(data %>% filter(treatment == 1 & post_treatment == 1) %>% pull(dins), data %>% filter(treatment == 1 & post_treatment == 1) %>% pull(W)) - weighted.mean(data %>% filter(treatment == 1 & post_treatment == 0) %>% pull(dins), data %>% filter(treatment == 1 & post_treatment == 0) %>% pull(W))

second_diff_W <- weighted.mean(data %>% filter(treatment == 0 & post_treatment == 1) %>% pull(dins), data %>% filter(treatment == 0 & post_treatment == 1) %>% pull(W)) - weighted.mean(data %>% filter(treatment == 0 & post_treatment == 0) %>% pull(dins), data %>% filter(treatment == 0 & post_treatment == 0) %>% pull(W))

(did_estimator_W <- first_diff_W - second_diff_W)
```


## Compute the DiD estimator for the ATT using regression with the treatment group, post-treatment timing dummies, as well as their interactions. Repeat the same exercise by using W as weights.

```{r}
# Without using W as weights:

regression <- data %>%
  estimatr::lm_robust(
    formula = dins ~ treatment + post_treatment + treatment*post_treatment
  )

(regression$coefficients[4])

# Using W as weights:

regression_W <- data %>%
  estimatr::lm_robust(
    formula = dins ~ treatment + post_treatment + treatment*post_treatment,
    weights = W
  )

(regression_W$coefficients[4])
```


## Compute the DiD estimator for the ATT using regression with state fixed effects, time fix effects, and the treatment dummy that takes value 1 if a unit is treated in the year 2014, and 0 otherwise. Repeat the same exercise by using W as weights.

```{r}
# Without using W as weights:

regression_FE <- data %>%
  estimatr::lm_robust(
    formula = dins ~ product,
    fixed_effects = ~ stfips + year
  )

(regression_FE$coefficients[1])

# Using W as weights:

regression_FE_W <- data %>%
  estimatr::lm_robust(
    formula = dins ~ product,
    fixed_effects = ~ stfips + year,
    weights = W
  )

(regression_FE_W$coefficients[1])
```


## Conduct statistical inference using your regression specifications. Make sure to cluster at the state level. What do the results suggest?

```{r}
# Specification from Question 6:

regression_clustered <- data %>%
  estimatr::lm_robust(
    formula = dins ~ treatment + post_treatment + treatment*post_treatment,
    cluster = stfips
  )

(regression_clustered$p.value[4])

# Specification from Question 7:

regression_FE_clustered <- data %>%
  estimatr::lm_robust(
    formula = dins ~ product,
    fixed_effects = ~ stfips + year,
    cluster = stfips
  )

(regression_FE_clustered$p.value[1])
```

Both p-values are the same, as we expected. We reject the null hypothesis of no treatment effect at the 1% significance level.